Neighbor-Based Decentralized Training Strategies for Multi-Agent Reinforcement Learning

### 2024-10-04 paper-2024-sac-marl-neighbouring-strategies.tex
fix final reward
more details on reward
add eval metrics description

First, we extended the \texttt{MultiAgentEnv} from RLlib to implement a custom cooperative environment 
  called \emph{collect the items} -- see \Cref{fig:env}.
%
This environment consists of a 2D euclidean space in which a variable number of \emph{items} and \emph{agents} 
 are randomly placed.
%
The goal of the agents is to cooperate to collect all the items in the smallest amount of time possible. 
%
In particular, each agent has an observation space defined by three main parameters, namely: 
\begin{enumerate*}[label=(\roman*)]
  \item number of closest agents that the agent is able to see (thus, forming a neighborhood);
  \item number of closest items that the agent is able to see;
  \item memory of past states, thus encoding also temporal information, helping each agent to understand 
    the evolution of the environment.
\end{enumerate*} 
%
Regarding the action space, each agent has $8$ possible moves, namely the possible direction on a grid 
  (up, down and diagonal).

% Reward

% Specify the chosen RL algorithm (DQN) and its hyperparameter configuration

% Define the evaluation metrics: average episode length and average reward

>>>>

\subsection{Environment} 
First, we extended the \texttt{MultiAgentEnv} from RLlib to implement a custom cooperative environment 
  called \emph{collect the items} -- see \Cref{fig:env}.
%
This environment consists of a 2D euclidean space in which a variable number of \emph{items} and \emph{agents} 
 are randomly placed.
%
The goal of the agents is to cooperate to collect all the items in the smallest amount of time possible. 
%
In particular, each agent has an observation space defined by three main parameters, namely: 
\begin{enumerate*}[label=(\roman*)]
  \item number of closest agents that the agent is able to see (thus, forming a neighborhood);
  \item number of closest items that the agent is able to see;
  \item memory of past states, thus encoding also temporal information, helping each agent to understand 
    the evolution of the environment.
\end{enumerate*} 
%
Regarding the action space, each agent has $8$ possible moves, namely the possible direction on a grid 
  (up, down and diagonal).

% Reward 
\subsection{Reward}
One of the most important aspects in MARL is how the reward signal that the agents use to 
  learn is defined.
%
In particular, we have defined a reward function composed of many subcomponents.

\paragraph{Collisions} Agents are penalized when their action area intersect. 
  \[ \psi = \begin{cases}
    0 & \text{if no intersections} \\
    -2 \cdot |I_a| & \text{otherwise}
  \end{cases} \]
  Where $I_a$ is the set of agents with intersecting action area.

\paragraph{Time step} Agents are penalized at each time step by a constant value $\sigma$ defined as
    an hyperparameter.
    \[ \zeta = - \sigma \]

\paragraph{Moving towards items} Each agent $a$ is rewarded if it moves towards one of the visible target items.
    \[ \xi = (max_{i \in V_{items}} d(p_a^{t+1}, p_i^{t+1}) - d(p_a^{t}, p_i^{t})) \cdot \delta\]
    Where $p_a^t$ is the position of the agent $a$ at time t (similarly for $p_i^{t}$ regarding items),
    $\delta$ is a scalar hyperparameter used to tune the importance of this component and
    $V_{items}$ is the set of visible items.
\paragraph{Moving away from agents} Each agent $a_i$ is rewarded if it increases the distance from its neighbors ($N_{a_i}$).
    \[ \chi = \frac{1}{2|N_{a_i}|} \cdot \sum_{a_j \in N_{a_i}} d(p_{a_i}^{t+1}, p_{a_j}^{t+1}) - d(p_{a_i}^{t}, p_{a_j}^{t}) \]

\paragraph{Collecting items} Each agent is rewarded for any collected visible items ($C_i$) is the set of collected items), 
    whether the item was collected by  the agent itself or by another agent, this strategy encourages agents 
    to develop cooperative behaviors.
    \[ \rho = 100 \cdot |C_i| \]

\paragraph{Final reward} The final reward is computed as the sum of all the subcomponents: 
  \[ R = \psi + \zeta + \xi + \chi + \rho \]

% Specify the chosen RL algorithm (DQN) and its hyperparameter configuration
\subsection{Learning algorithm}

% Define the evaluation metrics: average episode length and average reward
\subsection{Evaluation metrics} 
In our experiments, we employed two metrics to evaluate the performance of the learning
  agents and monitor the quality of the training process, namely: 
  \begin{enumerate*}[label=(\roman*)]
    \item \emph{Mean Episode Reward}; and
    \item \emph{Mean Episode Length}.
  \end{enumerate*}

\paragraph{Mean Episode Reward} 
The Mean Episode Reward measures the average cumulative reward the agent obtains during each episode, 
  serving as a direct indicator of how well each agent is performing in achieving its task.
%
Formally, for $N$ episodes, the mean reward $R_{mean}$ is computed as:
  \[ R_{mean} = \frac{1}{N} \sum_{i=1}^{N} R_i \]
  where $R_i$ is the total reward obtained in the $i$-th episode.

\paragraph{Mean Episode Length}
On the other hand, the Mean Episode Length represents the average number of time steps (or actions) 
  the agent takes per episode before reaching the terminal state. 
%
It provides insights into the agent's efficiency in completing tasks, especially in environments 
  where shorter episodes indicate better performance.
%
The mean episode length $L_{mean}$ is calculated as:
  \[ L_{mean} = \frac{1}{N} \sum_{i=1}^{N} L_i \]
  where $L_i$ is length of $i$-th episode.
